|

### 2023-02-02 paper/paper.Rmd
cleaned up paper.Rmd

# Iterate and save images
# Unconstrained orthonormalisation
Anew <- A
for (i in 1:3) {
  Anew[vchange,] <- A[vchange,]*(1-i*eps)
  exact <- Anew[vchange,]
  Anew <- tourr::orthonormalise(Anew)
  plot_basis(Anew, vchange, vcol="black")
  text(Anew[vchange, 1], Anew[vchange, 2], expression(V[m]), pos=3, col="orange")
  points(exact[1], exact[2], pch=16, col="orange")
}

# Iterate and save images
# Constrained orthonormalisation
#Anew <- A
#plot_basis(A, vchange, vcol="black")
#for (i in 1:3) {
#  Anew[vchange,] <- A[vchange,]*(1-i*eps)
#  exact <- Anew[vchange,]
#  Anew <- orthonormalise_frozen(Anew, vchange)
#  plot_basis(Anew, vchange, vcol="black")
#  points(exact[1], exact[2], pch=16, col="orange")
#}
```

@@

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data. The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
#ggscatmat(penguins_data, columns = 2:5, col="species") +
# library(paletteer) 
# rcartocolor::ag_Sunset
# rcartocolor::ag_GrnYl
# RColorBrewer::Spectral
# clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
# clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
# clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
#  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
#  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat + theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

>>>>

# Iterate and save images
# Unconstrained orthonormalisation
Anew <- A
for (i in 1:3) {
  Anew[vchange,] <- A[vchange,]*(1-i*eps)
  exact <- Anew[vchange,]
  Anew <- tourr::orthonormalise(Anew)
  plot_basis(Anew, vchange, vcol="black")
  text(Anew[vchange, 1], Anew[vchange, 2], expression(V[m]), pos=3, col="orange")
  points(exact[1], exact[2], pch=16, col="orange")
}
```

@@

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data. The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat + theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```
